% INTRO
\bibitem{barbosa2003displaced}
A. Barbosa, ``Displaced soundscapes: A survey of network systems for music and sonic art creation'',
\emph{Leonardo Music Journal},
pp. 53--59,2003,MIT Press
%
\bibitem{RottondiOverview}
C. Rottondi, C. Chafe, C. Allocchio, and A. Sarti
``An Overview on Networked Music Performance Technologies'', \emph{IEEE Access},4,2016
%
\bibitem{Lakiotakis}
E. Lakiotakis, C. Liaskos, X. Dimitropoulos, ``Improving networked music performance systems using application-network collaboration'', \emph{Concurrency and Computation: Practice and Experience},Wiley Online Library, 2018

\bibitem{RottondiFeature}
C. Rottondi, M. Buccoli, M. Zanoni, D. Garao,  G. Verticale and A. Sarti, ``Feature-Based Analysis of the Effects of Packet Delay on Networked Musical Interactions'', \emph{Journal of the Audio Engineering Society}, 63(11), 864-875.
% Chafe clapping experiments
\bibitem{Chafe3}
C. Chafe, J-P. Caceres, M. Gurevich, ``Effect of temporal separation on synchronization in rhythmic performance'',
\emph{Perception}, 39(7): 982-992, 2010
%
\bibitem{Chafe2}
C. Chafe and M. Gurevich, ``Network Time Delay and Ensemble Accuracy: Effects of Latency, Asymmetry'', \emph{Proc. of the AES 117th Conf.}, SF, 2004
%
\bibitem{Chafe1}
C. Chafe, M. Gurevich, et al. ``Effect of Time Delay on Ensemble Accuracy''
\emph{Proc. 2004 Intl. Soc. Musical Acoustics}, Nara, 2004
% BArbosa

\bibitem{Carot07networkmusic}
E. Car\^ot and C. Werner, ``Network music performance - problems, approaches and perspectives'',
\emph{International School of new Media, Institute of Telematics, University of Lübeck. Music in the Global Village - Conference} ,2007
%

\bibitem{gurevich2011ambisonic}
M. Gurevich, D. Donohoe and S. Bertet, ``Ambisonic Spatialization for Networked Music Performance'', \emph{International Community for Auditory Display}, 2011
% reverberation studies

\bibitem{carot2009towards}
A. Car{\^o}t, C. Werner and T. Fischinger, ``Towards a comprehensive cognitive analysis of delay-influenced rhythmical interaction'', \emph{ICMC}, 2009
%
\bibitem{FarnerReverb}
S. Farner, A. Solvang, A. S\aeb\o and Peter Svensson, A. Car{\^o}t, C. Werner and T. Fischinger, ``Ensemble hand-clapping experiments under the influence of delay and various acoustic
environments'',\emph{Journal of the Audio Engineering Society}, 57(12), pp 1028--1041, 2009
% LOLA

\bibitem{drioli2013networked}
C. Drioli, C. Allocchio, and N. Buso,
``Networked performances and natural interaction via LOLA: Low latency high quality A/V streaming system'',
\emph{Information Technologies for Performing Arts, Media Access, and Entertainment}, Springer, 2013
pp.240--250,
%Ultragrid
\bibitem{holub2006high}
P. Holub, L. Matyska, M. Li{\v{s}}ka, L. Hejtm{\'a}nek, J. Denemark, T. and Rebok, A. Hutanu, R. Paruchuri, J. Radil, and E. Hladk{\'a}
``High-definition multimedia for multiparty low-latency interactive communication'',
\emph{Future Generation Computer Systems},
22(8),
pp.856--861,
2006,
Elsevier
%Jacktrip
\bibitem{CaceresJacktrip}
J.-P. C\'{a}ceres, C. Chafe, ``JackTrip: Under the Hood of an Engine for Network Audio'', \emph{Proceedings of International Computer Music Conference}, Montreal, 2009. 

\bibitem{CIM2018}
S. Delle Monache, M. Buccoli, L. Comanducci, A. Sarti, G. Cospito, E. Pietrocola and F. Berbenni, ``Time is not on my side: network latency, presence and performance in remote music interaction.'' \emph{Proceedings of the XXII Colloquium on Musical Informatics (CIM)}, Udine,20-23 November, 2018.
%
\bibitem{Markovic2013} 
D. Markovic and F. Antonacci and A. Sarti and S. Tubaro, ``Soundfield Imaging in the Ray Space'',
\emph{IEEE Transactions on Audio, Speech, and Language Processing}, 
21(12), 
pp. 2493-2505, 2013
%
\bibitem{bianchi2016}
L. Bianchi, F. Antonacci, A. Sarti, S. Tubaro,
``Model-Based Acoustic Rendering based on Plane Wave Decomposition'',
\emph{Applied Acoustics},
Elsevier,
2016(104),
pp. 127--134
% Saccadic movements annotation
\bibitem{vandemoortele2018gazing}
S. Vandemoortele, K. Feyaerts, G. De Bi{\`e}vre, M. and Reybrouck, G. Br{\^o}ne, T. De Baets,
``Gazing at the partner in musical trios: a mobile eye-tracking study'',
\emph{Journal of Eye Movement Research}, 
Vol.11(2),2018
pp.6
% Biometric signal monitoring
\bibitem{Yoshie2009}
Yoshie, Michiko
and Kudo, Kazutoshi
and Murakoshi, Takayuki
and Ohtsuki, Tatsuyuki, ``Music performance anxiety in skilled pianists: effects of social-evaluative performance situation on subjective, autonomic, and electromyographic reactions'',
\emph{Experimental Brain Research},
vol.199(2),22 Aug.2009,pp.117
%
\bibitem{Woszczyk2009}
Wieslaw Woszczyk and William Martens, ``Evaluation of virtual acoustic stage support for musical performance'', \emph{Journal of the Acoustical Society of America}, Vol. 123(5), 2008, pp. 3089
%
\bibitem{Boucher15}
Matthew Boucher, David Pelegrin-Garcia, Bert Pluymers, and Wim Desmet, ``Auralization as a tool for evaluating an acoustical instrument'', \emph{Proc. of the Third Vienna Talk on Music Acoustics}, 16–19 Sept. 2015, University of Music and Performing Arts Vienna, pp. 19
%	
%	\bibitem{Cover}
%	\textit{(Example for books)} T.M.Cover and J.A. Thomas, \emph{Elements of Information Theory}. New York: Wiley, 1991.
%	
%	\bibitem{Dobrushin}
%	\textit{(Example for articles)} R.L. Dobrushin, ``Optimum information transmission through a channel with unknown parameters'',  \emph{Radiotech.Electron.}, vol.4, Dec.1959, pp. 1951-1956.
%	
%	\bibitem{Blachman}
%	\textit{(Example for articles)} N.M. Blachman, ``Communication as a game'', \emph{in Proc. WESCON Conf.}, Aug. 1957, pp. 61-66.
%	
%	\bibitem{IEEE}
%	\textit{(Example for web-links)} IEEE official website, Manuscript Templates for Conference Proceedings, Web: http://www.ieee.org/conferences\_events/conferences/publishing/templates.\newline html.
%	
%	\bibitem{Elissa}
%	\textit{(Example for unpublished references)} K. Elissa, ``Title of paper if known'', unpublished.
%	
%	\bibitem{Nicole}
%	\textit{(Example for references have been accepted for publication)} R. Nicole, ``Title of paper with only first word capitalized'', J. Name Stand. Abbrev., in press.
%	