
\bibitem{Lazzaro2001}
J.Lazzaro and J.Wawrzynek, ``A case for network musical performance'', in \emph{Proc. NOSSDAV Workshop}, Jun. 2001, pp. 157--166. 

\bibitem{IorwerthNMP2015}
M. Iorwerth, D. Moore and D. Knox, ``Challenges of using Networked Music Performance in education'', in  \emph{Proc. AES Conference}, Aug. 2015, pp. 157--166. 

\bibitem{MIDItoolbox}
T. Eerola and P. Toiviainen, \emph{MIDI Toolbox: MATLAB Tools for Music Research}, Finland: University of Jyv{\"a}skyl{\"a},  2004.

\bibitem{MOOCS}
N. Lushnikova, P. Chintakayala and A. Rodante, ``Massive Open Online Courses from Ivy League Universities: Benefits and Challenges for Students and Educators'' \emph{in proc. of the XI International Conference “Providing continuity of content in the system of stepwise graduate and postgraduate education}, Nov. 2012

\bibitem{Kolazi2013}
\c{S}. Kolozali and M. Barthet and G. Fazekas and M. Sandler, ``Automatic Ontology Generation for Musical Instruments Based on Audio Analysis'', in \emph{IEEE Transactions on Audio, Speech, and Language Processing}, Oct. 2013, pp. {2207--2220}
%	year={2013}, 
%	volume={21}, 
%	number={10}, 

\bibitem{camurri2010visual}
A. Camurri and T. Moeslund, Chap. ``Visual gesture recognition'' in \emph{Musical Gestures: Sound, Movement, and Meaning}, Edited by R. I. God{\o}y, and M. Leman, Routledge, 2010

\bibitem{Canclini2015}
A. Canclini, L. Mucci, F. Antonacci, A. Sarti, S. Tubaro, ``A Methodology for estimating the ratiationpattern of a violin during the performance'' \emph{in Proc. of the European Signal Processing Conference (EUSIPCO)}, 2015

\bibitem{Olmos2009}
A. Olmos, M. Brul\'{e}, N. Bouillot, M. Benovoy, J. Blum, H. Sun, N. W. Lund, and J. R. Cooperstock, ``Exploring the role of latency and orchestra placement on the networked performance of a distributed opera'' \emph{in 12th annual international workshop on presence}, 2009

\bibitem{Goto2010}
M. Goto, ``An Audio-based Real-time Beat Tracking System for Music With or Without Drum-sounds'' in \emph{Journal of New Music Research}, 30:2, pp. 159-171, 2010 

\bibitem{conductor2008}
A. Nijholt, D. Reidsma, R. Ebbers and M. ter Maat, ``The Virtual Conductor: Learning and Teaching about Music, Performing, and Conducting'', in \emph{proc. of the IEEE International Conference on Advanced Learning Technologies}, 2008

\bibitem{duffy2017new}
S. Duffy, and P. Healey, ``A new medium for remote music tuition'', in \emph{Journal of Music, Technology \& Education}, 2017, pp 5--27

\bibitem{uk5G}
UK 5G Innovation Network, Video: World's First 5G Distributed Music Concert, Web: https://uk5g.org/discover/read-articles/video-worlds-first-5g-distributed-music-concert/

\bibitem{CancliniAES}
A. Canclini, L. Mucci, F. Antonacci, A. Sarti and S. Tubaro, ``Estimation of the radiation pattern of a violin during the performance using plenacoustic methods'' \emph{in Audio Engineering Society Convention 138}

\bibitem{CancliniRoom}
A. Canclini, D. Marković, L. Bianchi, F. Antonacci, A. Sarti, S. Tubaro, ``A geometrical approach to room compensation for sound field rendering applications'' \emph{in Proc. of the European Signal Processing Conference (EUSIPCO)}, 2014

\bibitem{Bonacina2016}
L. Bonacina, A. Canclini, F. Antonacci, M. Marcon, A. Sarti, S. Tubaro, ``A low-cost solution to 3D pinna modeling for HRTF prediction'' in \emph{Proc. of the IEEE International Conference on Acoustics, Speech and Signal Processing (ICASSP)}, 2016

%	volume={10},
%	number={1},
%	pages={5--29},
%	year={2017},
%	publisher={Intellect}

%	m
%	\bibitem{Cover}
%	\textit{(Example for books)} T.M.Cover and J.A. Thomas, \emph{Elements of Information Theory}. New York: Wiley, 1991.
%	
%	\bibitem{Dobrushin}
%	\textit{(Example for articles)} R.L. Dobrushin, ``Optimum information transmission through a channel with unknown parameters'',  \emph{Radiotech.Electron.}, vol.4, Dec.1959, pp. 1951-1956.
%	
%	\bibitem{Blachman}
%	\textit{(Example for articles)} N.M. Blachman, ``Communication as a game'', \emph{in Proc. WESCON Conf.}, Aug. 1957, pp. 61-66.
%	
%	\bibitem{IEEE}
%	\textit{(Example for web-links)} IEEE official website, Manuscript Templates for Conference Proceedings, Web: http://www.ieee.org/conferences\_events/conferences/publishing/templates.\newline html.
%	
%	\bibitem{Elissa}
%	\textit{(Example for unpublished references)} K. Elissa, ``Title of paper if known'', unpublished.
%	
%	\bibitem{Nicole}
%	\textit{(Example for references have been accepted for publication)} R. Nicole, ``Title of paper with only first word capitalized'', J. Name Stand. Abbrev., in press.
%	